% --- SLIDE 26: Applications (I) ---
\begin{frame}{Applications of Quantum Entanglement (I)}

  \begin{block}{1. Quantum Computing}
    \begin{itemize}[<+->]
      \item Entangled \textbf{qubits} are the foundation of quantum computers.
      \item They allow for computations impossible for classical computers.
      \item Advances in cryptography, drug discovery, and materials science.
    \end{itemize}
  \end{block}

  \begin{alertblock}{2. Quantum Cryptography (QKD)}
    \begin{itemize}[<+->]
      \item Uses pairs of entangled photons to generate cryptographic keys.
      \item Offers \textbf{unconditional security} (guaranteed by the laws of physics).
      \item Any eavesdropping attempt disturbs the entanglement and is detected.
    \end{itemize}
  \end{alertblock}

\end{frame}

% --- SLIDE 27: Applications (II) ---
\begin{frame}{Applications of Quantum Entanglement (II)}

  \begin{block}{3. Quantum Communication and Networks}
    \begin{itemize}[<+->]
      \item Foundation for a future \textbf{quantum internet}.
      \item Enables hyper-secure communications and the interconnection of distributed quantum computers.
      \item Development of quantum repeaters for long distances.
    \end{itemize}
  \end{block}

  \begin{alertblock}{4. Quantum Metrology and Sensing}
    \begin{itemize}[<+->]
      \item Creates \textbf{sensors much more sensitive} than classical physics allows.
      \item Allows for measurements with precision beyond the ``standard quantum limit''.
      \item Applications in atomic clocks, GPS, weak-field detection, and microscopy.
    \end{itemize}
  \end{alertblock}

\end{frame}

% --- SLIDE 28: Applications (III) ---
\begin{frame}{Applications of Quantum Entanglement (III)}

  \begin{block}{5. Quantum Simulation and Teleportation}
    \begin{itemize}[<+->]
      \item \textbf{Quantum Simulation:} Modeling complex systems (molecules, materials) impossible for classical computers.
      \item \textbf{Quantum Teleportation:} Transferring quantum states (information) between distant particles.
    \end{itemize}
  \end{block}

\end{frame}
